---
title: "Data Cleaning - Homework #2"
author: "Bailey Jørgensen"
format: html
editor: source
execute: 
  warning: false
---

```{r}
# Load libraries
librarian::shelf(tidyverse, here, janitor)
```

```{r}
# Read in snowcover csv
snow <- read_csv(here("data", "processed", "snow_cover.csv"))

# Take a peek!
glimpse(snow)
```
```{r}
# Take a look at the counts for water cover
snow |> 
  count(Water_cover)
```

```{r}
# Take a look at the counts for land cover
snow |> 
  count(Land_cover)
```

*Both snow and land cover have interesting inputs, since as the "-" and ".". Plus, they are showing up as <chr>, even though they should be numeric. These are similar problems to the snow cover data, that we cleaned in class. So, lets do the same things here that we did there, and get this data all cleaned up and ready to use!*

```{r}
# Look at all of the funky symbols for water cover
snow |> 
  count(Water_cover) |> 
  filter(is.na(as.numeric(Water_cover)))
```
```{r}
# Look at all of the funky symbols for land cover
snow |> 
  count(Land_cover) |> 
  filter(is.na(as.numeric(Land_cover)))
```

*Lets try and figure out why all of these funky symbols got used*

```{r}
# View the rows that use the dots
snow |> 
  filter(Water_cover == ".") |> 
  View()
```

*Ah yes. As I expected. Those rows that use a "." seem to be the rows where all coverage types had a ".". In the snow_cover wrangling, we decided to convert all of these to NAs, as well as the -, n/a, and unk. From the looks of things, I want to do the same thing to the columns here. Imma get a lil fancy with it, to convert both water and land cover funky symbols to NAs, for both land and water*

```{r}
# Tidyverse will help us clean!
all_cover_fixed <- snow |> 
  mutate(across(
    c(Water_cover, Land_cover), 
    ~ if_else(.x %in% c(".", "-", "n/a", "unk"), NA_character_, .x)
  ))
```

```{r}
# Lets make sure that worked
all_cover_fixed |> 
  count(Water_cover) |> 
  filter(is.na(as.numeric(Water_cover)))
```
*All right! It worked! Now, the values that we assumed to be NA, are actually consistently represented as NA throughout the dataset.*

```{r}
# Now lets mutate to make the columns numeric
all_cover_fixed <- all_cover_fixed |> 
  mutate(Water_cover = as.numeric(Water_cover)) |> 
  mutate(Land_cover = as.numeric(Land_cover))
```

### Ok, we have dealt with the funky NA denotations. Now, since we are working with percentages, and assume that total cover should be a sum of these percentages, we want to make sure that our numbers are within an expected range (between 0 and 100)

```{r}
# Check for numbers over 100
all_cover_fixed |> 
  filter(Water_cover > 100)
```
```{r}
# Check for negative numbers
all_cover_fixed |> 
  filter(Land_cover <0)
```

*Ah oh, looks like we have a few numbers outside of our expected ranges. Unfortunately, I have to assume these are errors, and since I can't invent numbers to replace them, I will have to make these NAs as well.*

```{r}
# Turn negative and over 100 values to NA
all_cover_fixed <- snow |> 
  mutate(across(
    c(Water_cover, Land_cover),
    ~ if_else(.x > 100 | .x < 0, NA, .x)
  ))
```

```{r}
# Make sure it worked
all_cover_fixed |> 
  filter(Land_cover <0)
```

*Yay, it worked!*

*Now, lets take a look at the Total Cover column...why is it a character column, instead of numeric? Lets make sure there's nothing too crazy going on, before we recalcuate the total coverage*

```{r}
all_cover_fixed |> 
  count(Total_cover)
```
*Whoops, looks like there is ".", n/a, and "-" here too, as well as this crazy thing: <row r=\\64\\ spans=\\1:11\\ x14ac:dyDescent=\\0.2\\>. Time to do some cleaning of this column, and convert it to numeric*

```{r}
# Remove funky stuff, then convert to numeric
all_cover_fixed <- all_cover_fixed |> 
  mutate(Total_cover = if_else(Total_cover %in% c(".", "-", "n/a", "<row r=\\64\\ spans=\\1:11\\ x14ac:dyDescent=\\0.2\\>"), NA_character_, Total_cover)) |> 
  mutate(Total_cover = as.numeric(Total_cover))
```

```{r}
# Take a lil looksie
glimpse(all_cover_fixed)
```
*Ok, all is numeric now. Getting some numbers higher than 100 though, which doesn't make much sense from a percentage perspective. Odds are, this is because of the overly large numbers that we dealt with in our former cleaning. But, since that's all fixed, we can recalculate the total percentages as the sum of the three cover types, and it should be accurate!*

```{r}
all_cover_fixed <- all_cover_fixed |> 
  mutate(Total_cover_clean = rowSums(across(c(Snow_cover, 
                                              Water_cover, 
                                              Land_cover)), na.rm = TRUE))
```






